---
title: "R Basics with RStudio"
output: html_document
date: "2024-03-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Basic

Some shortcuts:

To clear the console:   Press `Ctrl` + `L` key combination.

To insert the assignment operator: Press `Alt` + `_` key combination.


## Load the `iris` dataset from the `datasets` library

```{r}
library(datasets)
data(iris)
head(iris)
```

Find the different species present in the data set use `unique()` command:

```{r}
unique(iris$Species)
```

## Install packages

Install the `GGally` package:

```{r eval=FALSE}
# Code to install "GGally" package
# This code will not executed be as the package has already been installed
install.packages("GGally", repos = "https://cran.r-project.org", type = "source")
```

## Creating Data Visualizations using `ggplot`

```{r}
data(mtcars)
head(mtcars, 5)
```

Type `?mtcars` to get information about the variables. This will print the information at the bottom right panel, on the `Help` tab.

```{r}
# load ggplot package
library(ggplot2)
```

Create a scatterplot of displacement `disp` and miles per gallon `mpg`:

```{r}
ggplot(aes(x=disp, y=mpg,), data=mtcars) + geom_point()
```

